Imminent range shifts in transboundary fish stocks challenge their management under climate change

### 2021-05-11 Scripts/Manuscript.Rmd
Fixed images links

---
title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
author: "SUBMITTED MANUSCRIPT. DO NOT SHARE."
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
subtitle: "Juliano Palacios-Abrantes*^1^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^"
output: bookdown::pdf_document2
reference_docx:  ../References/sci_template.docx
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
header-includes:
  - \usepackage{setspace}
  - \usepackage{caption}
  - \usepackage{float} # Use the 'float' package
  - \floatplacement{figure}{H} # Make every figure with caption = h
  - \captionsetup[figure]{name = Fig}
---

@@

```{r toeFigure1N, eval = F,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}

@@

The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig. \@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig. 2A). Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2 and \@ref(tab:toeTableS3) than almost any other region (Figs. \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)). In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts. Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig. \@ref(fig:toeFigure2)A).

@@

Furthermore, we developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock's abundance centroid and that of the EEZs that share such stock (Fig. \@ref(fig:toeFigure1)A)). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*). The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ. \@ref(equ:four)). We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged). In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig. \@ref(fig:toeFigureS8)). In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs. \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).

>>>>

---
title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
author: "SUBMITTED MANUSCRIPT. DO NOT SHARE."
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States"
subtitle: "Juliano Palacios-Abrantes*^1^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette C.C. Wabnitz^1,5^, and William W.L. Cheung^1^"
# output: bookdown::pdf_document2
output: bookdown::word_document2
reference_docx:  ../References/sci_template.docx
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
header-includes:
  - \usepackage{setspace}
  - \usepackage{caption}
  - \usepackage{float} # Use the 'float' package
  - \floatplacement{figure}{H} # Make every figure with caption = h
  - \captionsetup[figure]{name = Fig}
---

@@

```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}

@@

The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig. \@ref(fig:toeFigureS1)) and geographic region of the neighboring EEZs (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Table S3 and Fig. 2A). Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs. \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS1) and \@ref(fig:toeFigureS2)). In contrast, EEZs and stocks located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts. Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (Fig. \@ref(fig:toeFigure2)A).

@@

Furthermore, we developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock's abundance centroid and that of the EEZs that share such stock (Fig. \@ref(fig:toeFigure1)A)). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*). The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ. \@ref(eq:four)). We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged). In this case, the global average shift in the shared distribution of transboundary stocks by 2029 $\pm$ 27 years, a difference of about 7 years (Fig. \@ref(fig:toeFigureS8)). In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs. \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)).
